---
title: "STATS207 Project Analysis"
output: html_document
date: "2024-06-04"
---

```{r}
library(tidyverse)
library(zoo)
library(httr)
library(jsonlite)
library(astsa)
library(forecast)
library(tseries)
library(prophet)
```
# Data Fetching

```{r}
# Define API key
api_key <- "gymyk63by4IgrjFeoHZX606z9kwWdWhUhdO8zs9M"

get_data_for_year <- function(base_url, api_key, extra_params) {
  full_url <- paste0(base_url, "?api_key=", api_key, extra_params)
  response <- GET(full_url)
  
  if (status_code(response) == 200) {
    data <- content(response, "text", encoding = "UTF-8")
    json_data <- fromJSON(data)
    return(json_data$response$data)
  } else {
    return(NULL)
  }
}
base_url_gas <- "https://api.eia.gov/v2/natural-gas/cons/sum/data/"
extra_params_gas <- "&frequency=monthly&data[0]=value&facets[duoarea][]=NUS&facets[process][]=VCS&facets[process][]=VIN&facets[process][]=VRS&sort[0][column]=period&sort[0][direction]=desc&offset=0&length=5000"
gas_raw <- get_data_for_year(base_url_gas, api_key, extra_params_gas)

base_url_coal <- "https://api.eia.gov/v2/coal/consumption-and-quality/data/"
extra_params_coal <- "&frequency=quarterly&data[0]=consumption&facets[location][]=US&facets[sector][]=8&sort[0][column]=period&sort[0][direction]=desc&offset=0&length=5000"
coal_raw <- get_data_for_year(base_url_coal, api_key, extra_params_coal)

# U.S. Regular Gasoline All Sales/Deliveries by Prime Supplier (Thousand Gallons per Day)
base_url_pet <- "https://api.eia.gov/v2/petroleum/cons/prim/data/"
extra_params_pet <- "&frequency=monthly&data[0]=value&facets[duoarea][]=NUS&facets[product][]=EPMR&facets[series][]=C120000001&sort[0][column]=period&sort[0][direction]=desc&offset=0&length=5000"
petroleum_raw <- get_data_for_year(base_url_pet, api_key, extra_params_pet)

# possible extension to Consumption of Fuels for Electricity Generation and Useful Thermal Output
base_url_electricity <- "https://api.eia.gov/v2/electricity/retail/data/"

```

# Data Cleaning
## Petroleum
```{r}
petroleum <- petroleum_raw %>%
  select(period, value) %>%
  mutate(period = as.Date(as.yearmon(period)))%>%
  mutate(value = as.numeric(value))

ggplot(data = petroleum, aes( period, value )) + 
  geom_line() +
  theme_classic() +
  ylab("Gasoline Deliveries in the US (MMgal/day)")+
  xlab("Date")+
  geom_vline(xintercept = as.Date("2020-03-01"), color= "red")
```

## Coal
```{r}
coal <- coal_raw %>%
  select(period, consumption) %>%
  mutate(
    year = as.numeric(substring(period, 1, 4)),
    quarter = substring(period, 6, 7),
    month = case_when(
      quarter == "Q1" ~ 1,
      quarter == "Q2" ~ 4,
      quarter == "Q3" ~ 7,
      quarter == "Q4" ~ 10
    ),
    period = as.Date(paste(year, month, "01", sep = "-"))
  ) %>%
  select(-year, -quarter, -month) %>%
  mutate(consumption = as.numeric(consumption))

ggplot(data = coal, aes( period, consumption )) + 
  geom_line() +
  theme_classic() +
  ylab("Coal Commercial/Industrial Consumption in the US")+
  xlab("Date")+
  geom_vline(xintercept = as.Date("2020-03-01"), color= "red")+
  geom_rect(aes(xmin=as.Date("2000-01-01"), xmax=as.Date("2008-01-01"), ymin=0, ymax=Inf), alpha = 0.01)

coal <- coal %>%
  filter(period >= as.Date("2008-01-01"))

coal$consumption <- log(coal$consumption)

ggplot(data = coal, aes( period, consumption )) + 
  geom_line() +
  theme_classic() +
  ylab("log Transform")+
  xlab("Date")+
  geom_vline(xintercept = as.Date("2020-03-01"), color= "red")#+
  #geom_rect(aes(xmin=as.Date("2000-01-01"), xmax=as.Date("2008-01-01"), ymin=0, ymax=Inf), alpha = 0.01)

```

## Natural Gas
```{r}
gas <- gas_raw %>%
  mutate(period = as.Date(as.yearmon(period))) %>%
  filter(gas_raw$`series-description` == "Natural Gas Deliveries to Commercial Consumers (Including Vehicle Fuel through 1996) in the U.S. (MMcf)") %>%
  select(c(period, value))  %>%
  mutate(value = as.numeric(value))

ggplot(data = gas, aes( period, value )) + 
  geom_line() +
  theme_classic() +
  ylab("Natural Gas Delivered to Consumers in the US")+
  xlab("Date")+
  geom_vline(xintercept = as.Date("2020-03-01"), color= "red")
```
## Check completeness
```{r}
# Extract year and count occurrences per year
yearly_counts <- gas %>%
  mutate(year = format(period, "%Y")) %>%
  group_by(year) %>%
  summarise(count = n())

# Check if every year has 12 points
(yearly_counts$count == 12)

# Extract year and count occurrences per year
yearly_counts <- coal %>%
  mutate(year = format(period, "%Y")) %>%
  group_by(year) %>%
  summarise(count = n())

# Check if every year has 4 points
(yearly_counts$count == 4)

# Extract year and count occurrences per year
yearly_counts <- petroleum %>%
  mutate(year = format(period, "%Y")) %>%
  group_by(year) %>%
  summarise(count = n())

# Check if every year has 12 points
yearly_counts$count == 12
```

Yes, all years have 12 points, except for last year for gas and petroleum

# Data Analysis
## Restrict datasets to train only
```{r}
# separate train and test for each dataset
gas_full <- gas[order(gas$period), ]
gas_ts <- ts(gas_full$value, start = c(1973, 1), frequency = 12)
gas_train <- window(gas_ts, end = c(2020, 2))
gas_test <- window(gas_ts, start = c(2020, 3))

# previous df approach for gas
# gas_train <- gas_full %>%
#   filter(period < "2020-03-01")
# gas_test <- gas_full %>%
#   filter(period >= "2020-03-01")

# coal is quarterly
coal_full <- coal[order(coal$period), ]
coal_ts <- ts(coal_full$consumption, start = c(2008, 1), frequency = 4)
coal_train <- window(coal_ts, end = c(2020, 2))
coal_test <- window(coal_ts, start = c(2020, 3))

# previous df approach for coal
# coal_train <- coal_full %>%
#   filter(period < "2020-03-01")
# coal_test <- coal_full %>%
#   filter(period >= "2020-03-01")


petroleum_full <- petroleum[order(petroleum$period), ]
petroleum_ts <- ts(petroleum_full$value, start = c(1983, 1), frequency = 12)
petroleum_train <- window(petroleum_ts, end = c(2020, 2))
petroleum_test <- window(petroleum_ts, start = c(2020, 3))

# previous df approach for petroleum
# petroleum_train <- petroleum_full %>%
#   filter(period < "2020-03-01")
# petroleum_test <- petroleum_full %>% 
#   filter(period >= "2020-03-01")
```

## Fitting petroleum models
```{r petro_plot_og, echo=TRUE}

plot(petroleum_train)
acf2(petroleum_train) 
```
```{r petro_plot_piecewise, echo=TRUE}
petroleum_piecewise <- stl(petroleum_train, s.window = "periodic")
plot(petroleum_piecewise)
```
### Model 1: Manual SARIMA(4,2,1)(3,1,1)_12
```{r petro_plot_sarima, echo=TRUE}
petroleum_sarima <- sarima(petroleum_train, p = 12, d = 1, q = 2, P = 1, D = 1, Q = 1, S = 12)
petroleum_sarima

checkresiduals(petroleum_sarima$fit$residuals)

# check auto correlation of residuals
acf2(petroleum_sarima$fit$residuals)
```
```{r petro_sarima_forecast, echo=TRUE}
petroleum_forecast <- sarima.for(xdata = petroleum_train, n.ahead = 20, p = 12, d = 1, q = 2, P = 1, D = 1, Q = 1, S = 12, plot = TRUE)
```

### Model 2: Auto SARIMA
```{r petro_auto_sarima, echo=TRUE}
sarima_model <- auto.arima(petroleum_train, seasonal=TRUE, stepwise=FALSE, approximation=FALSE, trace=TRUE, allowdrift = TRUE)
sarima_forecast <- forecast(sarima_model, h=length(petroleum_test))
autoplot(sarima_forecast)
```

```{r petro_auto_arima_plot, echo=TRUE}
summary(sarima_model)
checkresiduals(sarima_model)

```

### Model 3: ETS
```{r petro_ets, echo=TRUE}
ets_model <- ets(petroleum_train, model="MAM", damped=TRUE)
ets_forecast <- forecast(ets_model, h=length(petroleum_test))
plot(ets_forecast)
```

```{r petro_ets_plot, echo=TRUE}
summary(ets_model)
plot(ets_model)
checkresiduals(ets_model)
```

### Model 4: TBATS
```{r petro_tbats, echo=TRUE}
tbats_model <- tbats(petroleum_train, seasonal.periods = 12)
tbats_forecast <- forecast(tbats_model, h=length(petroleum_test))
plot(tbats_forecast)
```

```{r petro_tbats_plot, echo=TRUE}
plot(tbats_model)
checkresiduals(tbats_model)
```
### Prophet Model
```{r petro_prophet, echo=TRUE}

petroleum_prophet_df <- petroleum_train %>%
  data.frame(ds = as.Date(time(petroleum_train)), y = petroleum_train)

# create holidays for prophet
df_holiday <- petroleum_prophet_df %>% mutate(holiday = ifelse(format(ds, "%m-%d") == "01-01", "newyear", 
                                        ifelse(format(ds, "%m-%d") == "07-04", "independence",
                                               ifelse(format(ds, "%m-%d") == "12-25", "christmas", NA))) ) %>% select(holiday, ds)

df_holiday <- df_holiday %>% filter(!is.na(holiday))

petroleum_prophet <- prophet(petroleum_prophet_df, holidays = df_holiday, yearly.seasonality = TRUE, weekly.seasonality = TRUE, daily.seasonality = TRUE, seasonality.mode = "multiplicative")

future <- make_future_dataframe(petroleum_prophet, periods = 20, freq = "month")
petroleum_forecast <- predict(petroleum_prophet, future)
plot(petroleum_prophet, petroleum_forecast)
```

```{r petro_prophet_plot, echo=TRUE}
petro_df_prophet <- head(petroleum_forecast, 446)
petro_df_prophet <- petro_df_prophet %>%
  mutate(period = ds, value = yhat, residuals = yhat - petroleum_train)

checkresiduals(petro_df_prophet$residuals)
```
### Cross Validation

```{r petro_cv, echo=TRUE, eval=FALSE}

# Create function to calculate RMSE
calc_rmse <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2))
}

# Load the data
ts_data <- petroleum_train

# Define the rolling window parameters
start_year <- 1990
end_year <- 2018
step_size <- 12 # Number of steps to forecast ahead
year = start_year
# Initialize vectors to store RMSE values
sarima_rmses <- c()
auto_sarima_rmses <- c()
ets_rmses <- c()
prophet_rmses <- c()
tbats_rmses <- c()

sarima_aics <- c()
auto_sarima_aics <- c()
ets_aics <- c()
tbats_aics <- c()

# Perform rolling window cross-validation
# Loop through the years
for (year in start_year:end_year) {
  
  # Define train and validation sets
  train_end <- c(year , 12)
  val_start <- c(year + 1, 1)
  val_end <- c(year + 1, 12)
  
  train <- window(ts_data, end=train_end)
  actual <- window(ts_data, start=val_start, end=val_end)
  
  # # 1. Fit SARIMA model
  # tryCatch({
  #   sarima_mod <- sarima(xdata = train, p = 12, d = 1, q = 2, P = 1, D = 1, Q = 1, S = 12)
  #   sarima_aics <- c(sarima_aics, sarima_mod$fit$aic)
  #   sarima_pred <- sarima.for(xdata = train, n.ahead = 12, p = 12, d = 1, q = 2, P = 1, D = 1, Q = 1, S = 12)$pred
  #   sarima_rmses <- c(sarima_rmses, calc_rmse(actual, sarima_pred))
  # }, error = function(e) {
  #   message("SARIMA model fitting failed: ", e)
  # })
  # 
  # # 1b. Fit auto SARIMA model
  # tryCatch({
  #   auto_sarima_model <- auto.arima(train, seasonal=TRUE, stepwise=FALSE, approximation=FALSE, trace=TRUE, allowdrift=TRUE)
  #   auto_sarima_aics <- c(auto_sarima_aics, auto_sarima_model$aic)
  #   auto_sarima_pred <- forecast(auto_sarima_model, h=12)$mean
  #   auto_sarima_rmses <- c(auto_sarima_rmses, calc_rmse(actual, auto_sarima_pred))
  # }, error = function(e) {
  #   message("Auto SARIMA model fitting failed: ", e)
  # })
  # 
  # # 2. Fit ETS model
  # tryCatch({
  #   ets_model <- ets(train, model="MAM", damped=TRUE)
  #   ets_aics <- c(ets_aics, ets_model$aic)
  #   ets_pred <- forecast(ets_model, h=12)$mean
  #   ets_rmses <- c(ets_rmses, calc_rmse(actual, ets_pred))
  # }, error = function(e) {
  #   message("ETS model fitting failed: ", e)
  # })
  # 
  # # 3. Fit Prophet model
  # tryCatch({
  #   # note, no AIC since Prophet is Bayesian
  #   prophet_df <- data.frame(ds=as.Date(time(train)), y=as.numeric(train))
  #   m <- prophet(prophet_df, yearly.seasonality=TRUE, weekly.seasonality=TRUE, daily.seasonality=TRUE, seasonality.mode="multiplicative", holidays=df_holiday)
  #   future <- make_future_dataframe(m, periods=12, freq="month")
  #   prophet_forecast <- predict(m, future)
  #   prophet_pred <- tail(prophet_forecast$yhat, 12)
  #   prophet_rmses <- c(prophet_rmses, calc_rmse(actual, prophet_pred))
  # }, error = function(e) {
  #   message("Prophet model fitting failed: ", e)
  # })
  
  # 4. Fit TBATS model
  tryCatch({
    tbats_model <- tbats(train)
    tbats_pred <- forecast(tbats_model, h=12)$mean
    tbats_rmses <- c(tbats_rmses, calc_rmse(actual, tbats_pred))
    tbats_aics <- c(tbats_aics, tbats_model$AIC)
  }, error = function(e) {
    message("TBATS model fitting failed: ", e)
  })
  
}

# Compare cross-validated performance
cat("SARIMA avg. RMSE:", mean(sarima_rmses), 
    "\nAuto SARIMA avg. RMSE:", mean(auto_sarima_rmses),
    "\nETS avg. RMSE:", mean(ets_rmses),
    "\nProphet avg. RMSE:", mean(prophet_rmses),
    "\nTBATS avg. RMSE:", mean(tbats_rmses))

cat("SARIMA avg. AIC:", mean(sarima_aics), 
    "\nAuto SARIMA avg. AIC:", mean(auto_sarima_aics),
    "\nETS avg. AIC:", mean(ets_aics),
    "\nTBATS avg. AIC:", mean(tbats_aics))
```


Plot sarima model along with the actual data
```{r petro_sarima_plot, echo=TRUE}
sarima_result <- sarima.for(xdata = ts_data, n.ahead = 25, p = 12, d = 1, q = 2, P = 1, D = 1, Q = 1, S = 12)
sarima_pred <- sarima_result$pred
sarima_upper = sarima_result$pred + 1.96 * sarima_result$se
sarima_lower = sarima_result$pred - 1.96 * sarima_result$se

# ggplot the forecasted values with the actual data

data_df <- data.frame(time=as.Date(time(petroleum_ts)), value=as.numeric(petroleum_ts))
  
pred_df <- data.frame(
  time = seq.Date(from = as.Date("2020-03-01"), by = "month", length.out = length(sarima_pred)),
  pred = sarima_pred,
  upper = sarima_upper,
  lower = sarima_lower
)
ggplot() +
  geom_line(data = data_df, aes(x = time, y = value), color = 'blue', size = 1) +
  geom_line(data = pred_df, aes(x = time, y = pred), color = 'red', size = 1) +
  geom_ribbon(data = pred_df, aes(x = time, ymin = lower, ymax = upper), alpha = 0.2, fill = 'red') +
  labs(title = 'Time Series with SARIMA Predictions', x = 'Time', y = 'Value') +
  theme_minimal()


```
## Fitting natural gas models
```{r ng_plot_og, echo=TRUE}

plot(gas_train)
acf2(gas_train) 
```
### Model 1: Manual SARIMA(5,1,1)(3,1,1)_12
```{r gas_plot_sarima, echo=TRUE}
gas_sarima <- sarima(gas_train, p = 5, d = 1, q = 1, P = 3, D = 1, Q = 1, S = 12)
gas_sarima

checkresiduals(gas_sarima$fit$residuals)
```
```{r gas_sarima_forecast, echo=TRUE}
gas_forecast <- sarima.for(xdata = gas_train, n.ahead = 20, p = 5, d = 1, q = 1, P = 3, D = 1, Q = 1, S = 12, plot = TRUE)
```

### Model 2: Auto SARIMA
```{r gas_auto_sarima, echo=TRUE}
sarima_model <- auto.arima(gas_train, seasonal=TRUE, stepwise=FALSE, approximation=FALSE, trace=TRUE, allowdrift = TRUE)
sarima_forecast <- forecast(sarima_model, h=length(gas_test))
autoplot(sarima_forecast)
```
```{r gas_auto_arima_plot, echo=TRUE}
summary(sarima_model)
checkresiduals(sarima_model)

```

### Model 3: ETS
```{r gas_ets, echo=TRUE}
ets_model <- ets(gas_train, model="MAM", damped=TRUE)
ets_forecast <- forecast(ets_model, h=length(gas_test))
plot(ets_forecast)
```

```{r gas_ets_plot, echo=TRUE}
summary(ets_model)
checkresiduals(ets_model)
```

### Model 4: TBATS
```{r gas_tbats, echo=TRUE}
tbats_model <- tbats(gas_train)
tbats_forecast <- forecast(tbats_model, h=length(gas_test))
plot(tbats_forecast)
```

```{r gas_tbats_plot, echo=TRUE}
summary(tbats_model)
checkresiduals(tbats_model)
```
### Prophet Model
```{r gas_prophet, echo=TRUE}

gas_prophet_df <- gas_train %>%
  data.frame(ds = as.Date(time(gas_train)), y = gas_train)

gas_prophet <- prophet(gas_prophet_df)

future <- make_future_dataframe(gas_prophet, periods = 20, freq = "month")
gas_forecast <- predict(gas_prophet, future)
plot(gas_prophet, gas_forecast)
```

```{r gas_prophet_plot, echo=TRUE}
gas_df_prophet <- head(gas_forecast, 230)
gas_df_prophet <- gas_df_prophet %>%
  mutate(period = ds, value = yhat, residuals = yhat - gas_train)

checkresiduals(gas_df_prophet$residuals)
```
### Cross Validation

```{r gas_cv, echo=TRUE, eval=FALSE}

# Create function to calculate RMSE
calc_rmse <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2))
}

# Load the data
ts_data <- gas_train

# Define the rolling window parameters
start_year <- 2008
end_year <- 2018 
step_size <- 12 # Number of steps to forecast ahead
year = start_year
# Initialize vectors to store RMSE values
sarima_rmses <- c()
auto_sarima_rmses <- c()
ets_rmses <- c()
prophet_rmses <- c()

sarima_aics <- c()
auto_sarima_aics <- c()
ets_aics <- c()

# Perform rolling window cross-validation
for (year in start_year:end_year) {
  
  # Define train and validation sets
  train_end <- c(year , 12)
  val_start <- c(year + 1, 1)
  val_end <- c(year + 1, 12)
  
  train <- window(ts_data, end=train_end)
  actual <- window(ts_data, start=val_start, end=val_end)
  
  # 1. Fit SARIMA model  
  sarima_mod <- sarima(xdata = train, p = 4, d = 1, q = 1,P = 3, D = 1, Q = 1, S = 12)
  sarima_aics <- c(sarima_aics, sarima_mod$fit$aic)
  sarima_pred <- sarima.for(xdata = train, n.ahead = 12, p = 4, d = 1, q = 1,P = 3, D = 1, Q = 1, S = 12)$pred
  sarima_rmses <- c(sarima_rmses, calc_rmse(actual, sarima_pred))
  
  # 1b. Fit auto SARIMA model
  auto_sarima_model <- auto.arima(train, seasonal=TRUE, stepwise=FALSE, approximation=FALSE, trace=TRUE, allowdrift=TRUE)
  auto_sarima_aics <- c(auto_sarima_aics, auto_sarima_model$aic)
  auto_sarima_pred <- forecast(auto_sarima_model, h=step_size)$mean
  auto_sarima_rmses <- c(auto_sarima_rmses, calc_rmse(actual, auto_sarima_pred))

  # 2. Fit ETS model
  ets_model <- ets(train, model="MAM", damped=TRUE)
  ets_aics <- c(ets_aics, ets_model$aic)
  ets_pred <- forecast(ets_model, h=step_size)$mean
  ets_rmses <- c(ets_rmses, calc_rmse(actual, ets_pred))
  
  # 3. Fit Prophet model
  # note, no AIC since Prophet is Bayesian
  prophet_df <- data.frame(ds=as.Date(time(train)), y=as.numeric(train))
  m <- prophet(prophet_df)
  future <- make_future_dataframe(m, periods=step_size, freq="month")
  prophet_forecast <- predict(m, future)
  prophet_pred <- tail(prophet_forecast$yhat, step_size)
  prophet_rmses <- c(prophet_rmses, calc_rmse(actual, prophet_pred))
}

# Compare cross-validated performance
cat("SARIMA avg. RMSE:", mean(sarima_rmses), 
    "\nAuto SARIMA avg. RMSE:", mean(auto_sarima_rmses),
    "\nETS avg. RMSE:", mean(ets_rmses),
    "\nProphet avg. RMSE:", mean(prophet_rmses))

cat("SARIMA avg. AIC:", mean(sarima_aics), # something is very strange here
    "\nAuto SARIMA avg. AIC:", mean(auto_sarima_aics),
    "\nETS avg. AIC:", mean(ets_aics))
```

## Fitting coal models
```{r petro_plot_og, echo=TRUE}

plot(coal_train)
acf2(coal_train) 
```
### Model 1: Manual SARIMA(2, 0, 0)(1, 1, 0)_4
```{r coal_plot_sarima, echo=TRUE}
coal_sarima <- sarima(coal_train, p = 2, d = 0, q = 0, P = 1, D = 1, Q = 0, S = 4)
coal_sarima

checkresiduals(coal_sarima$fit$residuals)
```
```{r coal_sarima_forecast, echo=TRUE}
coal_forecast <- sarima.for(xdata = coal_train, n.ahead = 20, p = 2, d = 0, q = 0, P = 1, D = 1, Q = 0, S = 4, plot = TRUE)
```

### Model 2: Auto SARIMA
```{r coal_auto_sarima, echo=TRUE}
sarima_model <- auto.arima(coal_train, seasonal=TRUE, stepwise=FALSE, approximation=FALSE, trace=TRUE, allowdrift = TRUE)
sarima_forecast <- forecast(sarima_model, h=length(coal_test))
autoplot(sarima_forecast)
```
```{r coal_auto_arima_plot, echo=TRUE}
summary(sarima_model)
checkresiduals(sarima_model)

```

### Model 3: ETS
```{r coal_ets, echo=TRUE}
ets_model <- ets(coal_train, model="MAM", damped=TRUE)
ets_forecast <- forecast(ets_model, h=length(coal_test))
plot(ets_forecast)
```

```{r coal_ets_plot, echo=TRUE}
summary(ets_model)
checkresiduals(ets_model)
```

### Model 4: TBATS
```{r coal_tbats, echo=TRUE}
tbats_model <- tbats(coal_train)
tbats_forecast <- forecast(tbats_model, h=length(coal_test))
plot(tbats_forecast)
```

```{r coal_tbats_plot, echo=TRUE}
summary(tbats_model)
checkresiduals(tbats_model)
```
### Prophet Model
```{r coal_prophet, echo=TRUE}

coal_prophet_df <- coal_train %>%
  data.frame(ds = as.Date(time(coal_train)), y = coal_train)

coal_prophet <- prophet(coal_prophet_df,
                        yearly.seasonality = TRUE)

future <- make_future_dataframe(coal_prophet, periods = 20, freq = "quarter")
coal_forecast <- predict(coal_prophet, future)
plot(coal_prophet, coal_forecast)
```

```{r coal_prophet_plot, echo=TRUE}
coal_df_prophet <- head(coal_forecast, 50)
coal_df_prophet <- coal_df_prophet %>%
  mutate(period = ds, value = yhat, residuals = yhat - coal_train)

checkresiduals(coal_df_prophet$residuals)
```
### Cross Validation

```{r coal_cv, echo=TRUE, eval=FALSE}

# Create function to calculate RMSE
calc_rmse <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2))
}

# Load the data
ts_data <- coal_train

# Define the rolling window parameters
start_year <- 2012
end_year <- 2018
step_size <- 4 # Number of steps to forecast ahead
year = start_year
# Initialize vectors to store RMSE values
sarima_rmses <- c()
auto_sarima_rmses <- c()
ets_rmses <- c()
prophet_rmses <- c()

sarima_aics <- c()
auto_sarima_aics <- c()
ets_aics <- c()

# Perform rolling window cross-validation
for (year in start_year:end_year) {
  
  # Define train and validation sets
  train_end <- c(year , 4)
  val_start <- c(year + 1, 1)
  val_end <- c(year + 1, 4)
  
  train <- window(ts_data, end=train_end)
  actual <- window(ts_data, start=val_start, end=val_end)
  
  # 1. Fit SARIMA model  
  sarima_mod <- sarima(xdata = train, p = 2, d = 0, q = 0, P = 1, D = 1, Q = 0, S = 4)
  sarima_aics <- c(sarima_aics, sarima_mod$fit$aic)
  sarima_pred <- sarima.for(xdata = train, n.ahead = 4, p = 2, d = 0, q = 0, P = 1, D = 1, Q = 0, S = 4)$pred
  sarima_rmses <- c(sarima_rmses, calc_rmse(actual, sarima_pred))
  
  # 1b. Fit auto SARIMA model
  auto_sarima_model <- auto.arima(train, seasonal=TRUE, stepwise=FALSE, approximation=FALSE, trace=TRUE, allowdrift=TRUE)
  auto_sarima_aics <- c(auto_sarima_aics, auto_sarima_model$aic)
  auto_sarima_pred <- forecast(auto_sarima_model, h=step_size)$mean
  auto_sarima_rmses <- c(auto_sarima_rmses, calc_rmse(actual, auto_sarima_pred))

  # 2. Fit ETS model
  ets_model <- ets(train, model="MAM", damped=TRUE)
  ets_aics <- c(ets_aics, ets_model$aic)
  ets_pred <- forecast(ets_model, h=step_size)$mean
  ets_rmses <- c(ets_rmses, calc_rmse(actual, ets_pred))
  
  # 3. Fit Prophet model
  # note, no AIC since Prophet is Bayesian
  prophet_df <- data.frame(ds=as.Date(time(train)), y=as.numeric(train))
  m <- prophet(prophet_df)
  future <- make_future_dataframe(m, periods=step_size, freq="quarter")
  prophet_forecast <- predict(m, future)
  prophet_pred <- tail(prophet_forecast$yhat, step_size)
  prophet_rmses <- c(prophet_rmses, calc_rmse(actual, prophet_pred))
}

# Compare cross-validated performance
cat("SARIMA avg. RMSE:", mean(sarima_rmses), 
    "\nAuto SARIMA avg. RMSE:", mean(auto_sarima_rmses),
    "\nETS avg. RMSE:", mean(ets_rmses),
    "\nProphet avg. RMSE:", mean(prophet_rmses))

cat("SARIMA avg. AIC:", mean(sarima_aics), # something is very strange here
    "\nAuto SARIMA avg. AIC:", mean(auto_sarima_aics),
    "\nETS avg. AIC:", mean(ets_aics))
```

Plot auto-arima (best model under cross validation)
```{r}
sarima_model <- auto.arima(coal_train, seasonal=TRUE, stepwise=FALSE, approximation=FALSE, trace=TRUE, allowdrift = TRUE)
sarima_forecast <- forecast(sarima_model, h=length(coal_test))
  
autoplot(sarima_forecast)+
  geom_line(data = coal, aes(x = c(1:nrow(coal))/4+2008-.25, y = rev(consumption)))
```